---
title: "Final Project Report"
author: "Mike,Craig,Belu"
date: "12/13/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

##Instructions (from canvas/grading rubric)
Length. No more than 5 pages of text, single spaced, excluding plots and tables. (Submit source code in a separate document.) As noted above, you may write less than 5 pages, provided that you satisfy the requirements for the project. Check the grading rubric on Canvas for specific grading criteria.

Report Format. Your final report should be written and formatted carefully. Consider your audience to be, hypothetically, the manager of a data science team. You should observe the best practices of statistical communication: use graphs when possible, labeling and explaining them, and interpret statistical results using language and quantities that non-statisticians can understand. (Remember: your results could get recycled for non-expert audiences.) Your report should: (1) introduce the problem, (2) describe the data and any cleaning you did, (3) explain your model in detail (how you developed it, and how it differs from and improves upon the model you used for the interim report), and (4) report model performance, including:

RMSE and R2 on the train set.
Estimated RMSE and R2 on the test set.
Your best Kaggle score (log RMSE) and rank.

```{r}
#MM - since code is hidden i figure we can put comments in the code chunks.  i was thinking for the format of the final report we can just use the outline from the grading rubric


```
## Introduction
Introduction. What is the problem you are working on? Briefly but thoroughly describe the Kaggle train/test format, and your ultimate objective in the project.

## Data modeling and cleaning
Data modeling and Cleaning. What sort of data modeling and cleaning decisions did you make? Did you create any variables? Did you refactor any variables? How did you deal with NAs?

## Model and model development
Model and Model development. Describe your model. Which variables did you end up using in your model, and why? Which statistical method(s) did you choose and why?

## Model performance
Model performance. How did your model perform on the train set? Report train set RMSE and R2. What is your estimate of the model’s performance on the test set using cross-validation? Report estimated RMSE and R2 on the test set, as well as the score you received back from your Kaggle submission.

## Statistical communication
Statistical communication. Are your figures labelled clearly, accompanied by explanatory captions, and referenced in the text? Likewise, are your tables titled and/or captioned appropriately and referenced in the text? Do the figures and tables chosen improve communication or are they superfluous?

##Overall quality
Overall quality. Are you writing in full sentences? Are there many gram- matical or spelling errors? Are you using the .Rmd or .Rnw formats correctly so that your final document is attractively formatted? Does the report look sharp, or are there issues with the compile from .Rmd such that there is lots of junk code getting printed to the screen?

## Leaderboard
Leaderboard rating. How good is your model compared to those that others reported.

